---
title: "COVID-19 EFFECTS ON FINANCIAL INCLUSION FOR MEXICO COMPARED TO THE US AND CANADA"
author: "Magdaly Santillanez"
date: "22.01.2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. After the covid-19, has the world changed in terms of financial inclusion?
- According to the World Bank (2022), financial inclusion is the cornerstone for development. After the COVID-19 pandemic there was an increase in accounts ownership and digital payments. However, it widened the gender gap worldwide. This study analyzes those trends to provide based evidence for public policy design.
- The primary data base is produced by the World Bank each three years, from 2011 to 2021, and has indicators in financial inclusion by country. This data is complemented with macroeconomic indicators to provide a wide view across time. 
- Mexico, United States and Canada are the main countries of the study. They formed a commercial alliance called USMCA, previous known as NAFTA. 

## 2. Retrieving the Global Financial Inclusion Index from the World Bank
- The database provides 800 indicators of financial inclusion across 160 countries. These indicators are summarized for the entire adult population and disaggregated by key demographic characteristics as gender, age, education, income, and are of residence (World Bank, 2024).  
- The dataset is [publicly accessible](https://databank.worldbank.org/databases). It was collected anonymously,and there are no ethical concerns regarding its use. 
- A potential limitation is that microdata, which allows more detail and disaggrergation, requires registration and can not be used to complement this research due to accessibility constraints. 

Before obtaining the data, let's load the required libraries:
```{r}
#Suppress Messages for avoiding unnecessary messages in the results
suppressPackageStartupMessages({
library(RSelenium)
library(tidyverse)
library(netstat)
library(rvest)
})
```

1. I will set up directories for this section: 

*Note: For full detail about the data management, please refer to Section 6 and to the READ ME file.*
```{r}
# 1. Consolidate directory creation
# Specify the project directory where is the script
project_dir <- getwd()

# Define new directories
#For non-processed data
raw_data <- file.path(project_dir, "raw_data") 
#Sub-folder for this query
findex_data <- file.path(raw_data, "findex_data")

# Create directories if they don't exist
if (!dir.exists(raw_data)) dir.create(raw_data)
if (!dir.exists(findex_data)) dir.create(findex_data)
```

2. Define URL for **primary** data source:
```{r}
url <- "https://databank.worldbank.org/source/global-financial-inclusion"
```

3. Set up Selenium and Browser:
```{r}
library(RSelenium)
library(netstat)  # for free_port()

# --- Standard Firefox session
fopts <- list("moz:firefoxOptions" = list(args = list()))  # add '--headless' if you want headless

rD <- rsDriver(
  browser    = "firefox",
  geckover   = "latest",     # GeckoDriver version (default is "latest")
  chromever  = NULL,         # don't download ChromeDriver
  iedrver    = NULL,         # (Windows only; skip)
  phantomver = NULL,         # <-- IMPORTANT: exclude PhantomJS to avoid the 402 call
  extraCapabilities = fopts,
  verbose    = FALSE,
  check      = FALSE,        # skip version checks to speed up / avoid extra downloads
  port       = free_port()
)

# Assign folder to browser
driver <- rD$client
# Navigate to the World Bank's Global Findex page
driver$navigate(url)
# Allow time for the page to load
Sys.sleep(2) 
```

4. Then, I will define a function for multiple clicks to avoid repetitive code. 

*Note*: I will include 'tryCatch' in case any error arises as random pop-up windows, which not always appear. 
```{r}
# Define the function
interaction_function <- function(driver, xpath, message_on_error = "Element was not found") {
  tryCatch({
    element <- driver$findElement(using = "xpath", value = xpath)
    element$clickElement()
    Sys.sleep(1)  
  }, error = function(e) {
    message(message_on_error)
  })
}
```

5. Define all Xpaths:
```{r}
# Define all type of XPaths: popups, filters, time, close indicator and download. 
xpaths <- list(
  popups = list(
    cookies = '//button[contains(text(), "No, thanks")]',
    feedback = '//*[@id="survey_img"]'
  ),
  filters = list(
    country = '//a[contains(text(), "Country")]',
    select_all_countries = '//a[contains(@class, "selectAll") and contains(@data-text, "Select All Country")]',
    add_series = '//a[contains(@class, "addMoreRHS") and contains(@title, "Add Series")]',
    indicators = c(
      '//*[@id="chk[FinDex_Series_DIM].[Topic].&[account.t.d]"]',
      '//*[@id="chk[FinDex_Series_DIM].[Topic].&[account.t.d.1]"]',
      '//*[@id="chk[FinDex_Series_DIM].[Topic].&[account.t.d.2]"]',
      '//*[@id="chk[FinDex_Series_DIM].[Topic].&[borrow.any]"]',
      '//*[@id="chk[FinDex_Series_DIM].[Topic].&[g20.made.t.d]"]',
      '//*[@id="chk[FinDex_Series_DIM].[Topic].&[save.any]"]'
    ),
    add_time = '//*[@id="addTimeBtn"]',
    select_all_time = '//a[contains(@class, "btn selectAll btn-default") and contains(@data-text, "Select All Time")]',
    close_popup = '//button[@title="Close"]'
  ),
  download = list(
    dropdown = '//a[@title="Download options"]',
    csv = '//a[@title="Download CSV Format"]'
  )
)
```

6. Execute all previously defined interactions
```{r}
# Handle pop-ups
interaction_function(driver, xpaths$popups$cookies, "The cookie pop-up did not appear, I will proceed with next steps")
interaction_function(driver, xpaths$popups$feedback, "No feedback pop-up found, I will proceed with next steps")

# Interact with filters
interaction_function(driver, xpaths$filters$country)               # Select "Country"
interaction_function(driver, xpaths$filters$select_all_countries)  # Select "All Countries"
interaction_function(driver, xpaths$filters$add_series)            # For "Add Series"

# Loop through the indicators and click on them
for (xpath in xpaths$filters$indicators) {
  interaction_function(driver, xpath, sprintf("Error interacting with indicator: %s", xpath))
}

# Add time and select all time periods
interaction_function(driver, xpaths$filters$add_time, "'Add Time' button not found")
interaction_function(driver, xpaths$filters$select_all_time, "Select 'All Time' button not found")

# Close the Pop-Up
interaction_function(driver, xpaths$filters$close_popup, "'Close Pop-Up' button not found")

# File download
interaction_function(driver, xpaths$download$dropdown, "'Download dropdown' not found")
Sys.sleep(3)  # Added delay before clicking download 
interaction_function(driver, xpaths$download$csv, "'CSV download' option not found")
Sys.sleep(5)  # Added delay for download
message("File download has initiated. The directory is: ", findex_data)
```

7. Stop Selenium
```{r}
driver$close()
rD$server$stop()
```

8. Un-zip the file and change name for a more descriptive indicator
```{r}
#Specify file location
zip_file <- file.path(findex_data, "P_Data_Extract_From_Global_Financial_Inclusion.zip")
# Unzip the file 
unzip( zip_file, exdir = findex_data)

#Replace name for a clearer descriptor
downloaded_files <- list.files(findex_data, full.names = TRUE)

  # 1. Rename the first file to "findex_data.csv"
  file.rename(downloaded_files[1], file.path(findex_data, "findex_data.csv"))
  
  # 2. Rename the second file to "findex_series.csv"
  file.rename(downloaded_files[2], file.path(findex_data, "findex_series.csv"))

#Read the file:
findex_data_file <- file.path(findex_data, "findex_data.csv")
findex_indicators <- read_csv(findex_data_file, show_col_types = FALSE) 
```

## 3. Collecting main macro economic indicators and geo-spatial data

Now, I will collect:

  A. The **macro indicators** will provide critical context for understanding the financial behavior, inclusion patters and    changes for USMCA countries. While,
  
  B. The **geo-spatial information** will allow to graphically understand the main changes worldwide before and after the      pandemic.

1. Let's load the API URL for the **macro indicator's** query by country and create folder: 

```{r}
#URL
country_url <- "https://api.worldbank.org/v2/country/"

# Path for new 'API' data inside 'raw_data' folder
API_data <- file.path(raw_data, "API_data")

# Check if the 'API data' folder exists, and create it if it doesn't
if (dir.exists(API_data)) {} else {
  dir.create(API_data)
}
```

2. Load the required libraries: 
```{r}
#Add suppress start up messages for unnecessary content
suppressPackageStartupMessages ({
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
})
```

For the query, I will define the list of indicators as inflation, GDP, Population, Education Expenditure, Poverty Ratio, Interest Rate, and Unemployment. The objective is to compare USMCA  financial inclusion performance in the pandemics, coupled with main macro-economic indicators. 

According to the [API documentation](https://documents.worldbank.org/en/publication/documents-reports/api), it is not possible to consult more than one indicator at the same time. But it is possible to do it for multiple countries for the same indicator. Then, an optimal way to reduce the number of consults was to narrow the number of indicators while querying for all the countries at the same time. 

3. Define the parameters, countries and time scope: 
```{r}

# Countries
countries <- data.frame(
  iso3 = c("MEX", "USA", "CAN"),  # ISO3 codes
  name = c("Mexico", "United States", "Canada")  # Country names
)

#Parameters
country_codes <- paste(countries$iso3, collapse = ";")  # Combine ISO3 codes for identify countries

indicators <- c(
  "FP.CPI.TOTL.ZG",        # Inflation
  "NY.GDP.MKTP.CD",        # GDP 
  "SP.POP.TOTL",           # Population
  "SE.XPD.TOTL.GD.ZS",     # Expenditure on education
  "SI.POV.DDAY",           # Poverty ratio
  "FR.INR.RINR",           # Interest rate 
  "SL.UEM.TOTL.ZS"         # Unemployment
)

#Time
start_year <- 2011
end_year <- 2024
```
*Note: The indicators and countries codes were obtained from the API documentation.*

4. Now I will obtain details for indicators: I will consult the indicators API to create a catalog. Note this API is different from the previous one.
```{r}
indicators_url <- "https://api.worldbank.org/v2/indicator/"
```

5. Get the data:
```{r}
get_indicators <- function(indicator) {
  url <- paste0(indicators_url, indicator, "?format=json")
  response <- fromJSON(content(GET(url), as = "text"))
  if (length(response) < 2 || is.null(response[[2]])) return(NULL)
  response[[2]] %>%
    as.data.frame() %>%
    select(id, name, sourceNote) %>%
    rename(Series_Code = id, Indicator = name, Description = sourceNote)
}

#Establish file path
indicators_catalog <- bind_rows(lapply(indicators, get_indicators))
```

6. Then, I will design the query for the World Bank API for a single indicator and multiple countries to reduce as much as possible the number of consults.
```{r}
# Function to collect data for a single indicator for multiple countries
indicator_data <- function(indicator, country_codes, start_year, end_year) {
  url <- paste0(country_url, country_codes, "/indicator/", indicator, "?format=json&date=", start_year, ":", end_year)
  response <- GET(url)
  
  if (status_code(response) != 200) {
    stop("Failed to collect data for indicator: ", indicator)
  }
  
  data <- fromJSON(content(response, as = "text"))
  
  # Check if data is valid
  if (length(data) < 2 || is.null(data[[2]])) {
    warning("No data available for indicator: ", indicator)
    return(NULL)
  }
  
  # Extract the data and add the description
  df <- data[[2]] %>%
    select(date, value, countryiso3code) %>%
    mutate(
      Series_Code = indicator 
    )
  return(df)
}

# Collect data for all indicators and include descriptions
all_indicator_data <- lapply(indicators, function(ind) {
  indicator_data(ind, country_codes, start_year, end_year)
})

#USMCA raw
USMCA_indicators_raw <- bind_rows(all_indicator_data)
```


Now, let's download the **geospatial data**. This data is publicly available and comes from [North American Cartographic Information Society (NACIS)](https://www.naturalearthdata.com), which supports the sharing of geographic and spatial data resources. 

1. Define the URL for downloading the shapefile 
```{r}
geom.data <- "https://naciscdn.org/naturalearth/10m/cultural/ne_10m_admin_0_countries.zip"
```

2. Define the folder and the path for downloading it: 
```{r}
# Path for new 'shapefile' folder inside 'raw_data' folder
shape_data <- file.path(raw_data, "shape_data")

# Check if the 'shapefile' folder exists, and create it if it doesn't
if (dir.exists(shape_data)) {} else {
  dir.create(shape_data)
}

# File path to save the downloaded file
geom.file <- file.path(shape_data, basename(geom.data))
```

3. Un-zip the file and store it in the folder: 
```{r}
# Check if the file already exists
if (!file.exists(geom.file)) {
  message("Downloading the shape data...")
#Download and unzip the file: 
  download.file(geom.data, geom.file, mode = "wb") 
  unzip(geom.file, exdir = shape_data)
  
# Confirmation message
  message("File downloaded successfully to: ", geom.file)
} else {
  message("File already exists at: ", geom.file)
}
#Read shp file:
suppressPackageStartupMessages (library(sf))
world_shp <- st_read(file.path(shape_data, "ne_10m_admin_0_countries.shp"))
```

## 4. Transformation of the data for the analysis 

1. Create directory 
```{r}
# Define directory
processed_data <- file.path(project_dir, "procesed_data")

# Create directory if it does not exist
if (!dir.exists(processed_data)) dir.create(processed_data)
```

2. Homologate name of variables, data type, and not required data.
```{r}
library(tidyverse)
library(dplyr)

# 1. Findex indicators to long version: 
findex_indicators_long <- findex_indicators %>%
  pivot_longer(
    cols = starts_with("20"),  # Automatically selects columns starting with '20'
    names_to = "Year",
    values_to = "Value"
  ) %>%
  mutate(Year = as.numeric(str_extract(Year, "\\d{4}")), #Convert year to numeric
          Value = as.numeric(Value)) %>% #Convert Value to numeric  
  rename(
    Country = `Country Name`,
    Country_Code = `Country Code`,
    Indicator = `Series Name`,
    Series_Code = `Series Code`
  ) %>% select (all_of(order(colnames(.))))


# 2. USMCA Indicators: Combine, transform, and reshape indicators data
USMCA_indicators <- bind_rows(all_indicator_data) %>%
  mutate(Year = as.numeric(date)) %>%  
  select(-date) %>%
  rename(
    Value = value,
    Country_Code = countryiso3code
  ) %>%
  arrange(Country_Code, Year) %>%
  left_join(countries, by = c("Country_Code" = "iso3")) %>%
  rename(Country = name)%>% 
  left_join(indicators_catalog %>% select (Series_Code, Indicator),
  by = c("Series_Code" = "Series_Code")) %>%
  select (all_of(order(colnames(.))))

# 3. Review column names and same type
colnames(findex_indicators_long)
colnames(USMCA_indicators)

str(findex_indicators_long)
str(USMCA_indicators)

# 4. Combine and save: 
all_indicators <- bind_rows(findex_indicators_long, USMCA_indicators)
```

3. Ensure data is tidy tabular data: The following information demonstrates that each column represents a variable, each row represents an observation and each cell contains a single value.
```{r}
# Verify the structure of the dataset
head(all_indicators, n=3)
```

4. Now, I will proceed with doing 5 transformations to the all indicators data base:
```{r}
# 1. Calculate the GDP growth
gdp_growth <- all_indicators %>%
  filter(Series_Code == "NY.GDP.MKTP.CD") %>% # Select only GDP-related rows
  group_by(Country) %>%
  mutate(
    GDP_Growth = (Value - lag(Value)) / lag(Value) * 100  # Calculate GDP growth
  ) %>%
  ungroup()

# 2. Create a categorical variable for pre and post COVID-19
all_indicators <- all_indicators %>%
  mutate(Period = ifelse(Year < 2020, "Pre-COVID", "Post-COVID"))

# 3. Calculate the amount of government expenditure in education
educ_exp <- all_indicators%>%
  filter(Series_Code %in% c("SE.XPD.TOTL.GD.ZS", "NY.GDP.MKTP.CD")) %>%
  select(Country, Year, Series_Code, Value) %>%
  pivot_wider(names_from = Series_Code, values_from = Value) %>%
  mutate(
    Education_Expenditure = SE.XPD.TOTL.GD.ZS / 100 * NY.GDP.MKTP.CD  # Calculating expenditure
  )

# 4. Calculate gender gap: difference between male and female account ownership
gender_gap <- all_indicators %>%
  filter(Series_Code %in% c("account.t.d.1", "account.t.d.2")) %>% # Filter for Female and Male
  mutate(Value = as.numeric(Value)) %>% # Ensure Value is numeric
  group_by(Country, Year) %>%
  summarize(
    Female = sum(Value[Series_Code == "account.t.d.1"], na.rm = TRUE),
    Male = sum(Value[Series_Code == "account.t.d.2"], na.rm = TRUE),
    Gender_Gap = Male - Female, # Calculate the gap
    .groups = "drop" ) %>%       
  mutate(
    Female = na_if(Female, 0),
    Male = na_if(Male, 0),
    Gender_Gap = na_if(Gender_Gap, 0)) # Replace zero values with NA

# 5. Normalize inflation rates
inflation_normalized <- all_indicators %>%
  filter(Series_Code == "FP.CPI.TOTL.ZG") %>%  # Select indicator
  group_by(Country) %>%
  mutate(
    Inflation_Z_Score = (Value - mean(Value, na.rm = TRUE)) / sd(Value, na.rm = TRUE)  # Formula for Z-score 
  ) %>%
  ungroup()
```

## 5. What can the data tell us? 

Let's analyze the global situation and then, move to the three countries object of the study: 
 
1. Load libraries: 
```{r}
# Load necessary libraries
suppressPackageStartupMessages({
  library(ggplot2)
  library(tidyverse)
  library(sf)
  library(terra)
  library(readxl)
  library(viridis)
})
```

And create the folder for storing the visual outputs: 
```{r}
# Define directory
graphics_output <- file.path(project_dir, "outputs")
# Create directory if it does not exist
if (!dir.exists(graphics_output)) dir.create(graphics_output)
```

2. Prepare data:
```{r}
# Filter data for the indicator 'Account (% age 15+)' in relevant years
findex_account_map <- findex_indicators_long %>%
  filter(Series_Code %in% c("account.t.d"), 
         Year %in% c(2017, 2021, 2022)) %>%
  drop_na(Value)

# Merge the World Bank data with the shapefile
merged_df <- world_shp %>%
  left_join(findex_account_map, by = c("ADM0_A3" = "Country_Code"))
```

3. Create plot: 

- The selected colors are not from a colorblind friendly palette. But, the semaphore design highlights how countries improved their access to bank accounts by changing from reddish to greener tones. 

```{r}
# Function to generate the plot
generate_plot <- function(data, year_range) {
  data %>%
    filter(Year %in% year_range) %>%
    ggplot() +
    geom_sf(aes(fill = Value), color = "white", linewidth = 0.2) +  # Map aesthetics
    scale_fill_gradientn(
      colors = c("#A50026", "#FDAE61", "#1A9850"),  # Custom gradient colors
      name = "% Account Access (age 15+)",
      na.value = "black"  # Color for non-data
    ) +
    labs(
      title = paste("Global Financial Inclusion:", paste(year_range, collapse = ", ")),
      subtitle = "Percentage of Adults with Bank Account Access",
      caption = "Source: World Bank Global Findex, 2021."
    ) +
    theme_minimal(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 10)),
      plot.caption = element_text(size = 10, color = "darkgrey", hjust = 1),
      legend.position = "top",
      legend.key.width = unit(2, "cm"),
      legend.key.height = unit(0.4, "cm"),
      legend.text = element_text(size = 8),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank()
    )
}

# Generate and display plots for 2017 and 2021 (or 2022)
plot_2017 <- generate_plot(merged_df, 2017)
plot_2021 <- generate_plot(merged_df, c(2021, 2022))

# Print the plots
print(plot_2017)
print(plot_2021)

```

Now, I will analyze for the USMCA countries: 

- One limitation is that I do not have enough information about the interest rate. Then, I will use the inflation rate as a proxy to analyze how much people paid for borrowing money. 
```{r}
# Omit NA values and ensure 'Year' is included
data_filtered_2 <- all_indicators %>%
  filter(Country_Code %in% c("MEX", "CAN", "USA")) %>%
  filter(Series_Code %in% c("borrow.any", "FP.CPI.TOTL.ZG", "FR.INR.RINR")) %>%
  drop_na(Value) # Remove rows with NA in specified columns

# Split data for bars and lines
data_borrow <- data_filtered_2 %>%
  filter(Series_Code == "borrow.any")

data_inflation <- data_filtered_2 %>%
  filter(Series_Code == "FP.CPI.TOTL.ZG")

data_interest <- data_filtered_2 %>%
  filter(Series_Code == "FR.INR.RINR")

borrow_inflation <- ggplot() +
  # Bars for Borrowing Behavior
  geom_bar(
    data = data_borrow,
    aes(x = Year, y = Value, fill = Country),
    stat = "identity",
    alpha = 0.6,
    position = "dodge"
  ) +
  # Lines for Inflation with linetypes and markers
  geom_line(
    data = data_inflation,
    aes(x = Year, y = Value * 10, color = Country, group = Country, linetype = Country),
    linewidth = 0.8
  ) +
  geom_point(
    data = data_inflation,
    aes(x = Year, y = Value * 10, color = Country, shape = Country),
    size = 0.8
  ) +
  # Annotations for events (e.g., COVID-19)
  annotate("rect", xmin = 2020, xmax = 2021, ymin = -Inf, ymax = Inf, 
           alpha = 0.1, fill = "grey") +
  annotate("text", x = 2020.5, y = 50, label = "COVID-19 Impact", 
           hjust = 0.5, color = "black", size = 4) +
  # Labels and Scales
  labs(
    title = "Borrowing Behavior and Inflation Over Time (USMCA Countries)",
    subtitle = "Overlapping Financial Inclusion (Bars) and Inflation (Lines)",
    x = "Year",
    y = "Borrowing Behavior (% age 15+)",
    caption = "Source: World Bank Global Findex and World Development Indicators, 2021.",
    fill = "Borrowing (Bars)",
    color = "Inflation (Lines)",
    linetype = "Country",
    shape = "Country"
  ) +
  # Y-axis with secondary scale
  scale_y_continuous(
    name = "Borrowing Behavior (% age 15+)",
    sec.axis = sec_axis(~ . / 10, name = "Inflation Rate (%)") # Adjust inflation scaling
  ) +
  # Colors for bars and lines
  scale_fill_manual(
    values = c(
      "Mexico" = "#A50026",    
      "Canada" = "#FDAE61",     
      "United States" = "#1A9850"
    )
  ) +
  scale_color_manual(
    values = c(
      "Mexico" = "#A50026",    
      "Canada" = "#FDAE61",     
      "United States" = "#1A9850"
    )
  ) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  scale_shape_manual(values = c(16, 17, 18)) + # Circle, triangle, square
  # Theme for Clarity
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 10, color = "darkgrey", hjust = 1),
    legend.position = "top",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    axis.title.y.right = element_text(color = "black"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    panel.grid.major.y = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    plot.margin = margin(20, 20, 20, 20)
  )

print(borrow_inflation)

```

- How was the gender gap in account ownership across time?
```{r}
# Filter data for USMCA
filtered_data <- gender_gap %>%
  filter(Country %in% c("Mexico", "Canada", "United States") & !is.na(Gender_Gap))

account_trend <- ggplot(filtered_data, aes(x = Year, y = Gender_Gap, color = Country, group = Country)) +
  # Lines and Points for better visualization
  geom_line(linewidth = 1, aes(linetype = Country)) +
  geom_point(size = 2, aes(shape = Country)) +
  
  annotate("rect", xmin = 2020, xmax = 2021, ymin = -Inf, ymax = Inf, alpha = 0.2, fill = "grey") +
  annotate("text", x = 2021, y = 5, label = "COVID-19 Impact", hjust = 0.5, vjust = 1.5, color = "black", size = 4) +
  
  # Colors
  scale_color_manual(
    name = "Country",
    values = c(
      "Mexico" = "#A50026",    
      "Canada" = "#FDAE61",     
      "United States" = "#1A9850"
    )
  ) +
  
  # Axis and Gridlines
  scale_x_continuous(
    breaks = c(2011, 2014, 2017, 2021, 2022),
    labels = c("2011", "2014", "2017", "2021", "2022")
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.1))  # Add space above/below points
  ) +
  
  # Labels
  labs(
    title = "Gender Gap in Account Ownership (USMCA Countries)",
    subtitle = "Percentage difference in account ownership between men and women (2011-2022)",
    x = "Year",
    y = "Gender Gap (%)",
    caption = "Source: World Bank Global Findex, 2021."
  ) +
  
  # Theme minimal
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 10, color = "darkgrey", hjust = 1, margin = margin(t = 10)),
    legend.position = "top",
    legend.title = element_text(face = "bold", size = 11),
    legend.text = element_text(size = 10),
    legend.key = element_blank(),
    legend.spacing.x = unit(0.3, "cm"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    panel.grid.major.y = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(20, 20, 20, 20)
  )

print(account_trend)

```

Since 2017 the gender gap for Mexico has widened, being worse in 2021.

-Finally, how did savings, borrowing and digital payments perform? Is there a possible correlation which could be further analyzed? 
```{r}
# Renaming indicators for the graph
data_filtered_3 <- all_indicators %>%
  filter(Country_Code %in% c("MEX", "CAN", "USA")) %>%
  filter(Series_Code %in% c("g20.made.t.d", "save.any", "borrow.any")) %>%
  drop_na(Value) %>%
  mutate(Series_Code = case_when(
    Series_Code == "g20.made.t.d" ~ "Digital Payments",
    Series_Code == "save.any" ~ "Savings",
    Series_Code == "borrow.any" ~ "Borrowing",
    TRUE ~ Series_Code
  ))

# Creating the plot
main_trends <- ggplot(data_filtered_3, aes(x = Year, y = Value, group = interaction(Country_Code, Series_Code), 
                            color = Country_Code, linetype = Series_Code)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8, aes(shape = Series_Code)) +
  
  # Identifying COVID-19 period
  annotate("rect", xmin = 2020, xmax = 2021, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.2) +
  annotate("text", x = 2020.5, y = max(data_filtered_3$Value, na.rm = TRUE), 
           label = "COVID-19", hjust = 0.5, color = "black", size = 3) +
  

  facet_wrap(~Country_Code, nrow = 1) +
  
  # Theme colors
  scale_color_manual(
    values = c("MEX" = "#A50026", "CAN" = "#FDAE61", "USA" = "#1A9850"),
    name = "Country"
  ) +
  scale_shape_manual(
    values = c("Digital Payments" = 16, "Savings" = 17, "Borrowing" = 18),  
    name = "Indicator"
  ) +
  scale_linetype_manual(
    values = c("Digital Payments" = "solid", "Savings" = "dashed", "Borrowing" = "dotted"),  # Corrected levels
    name = "Indicator"
  ) +
  
  # Labels and titles
  labs(
    title = "Trends in Digital Payments, Savings, and Borrowing (USMCA Countries)",
    subtitle = "Data from 2011 to 2022 with a focus on the COVID-19 period.",
    x = "Year",
    y = "Indicator Value",
    caption = "Source: World Bank Global Findex, 2021."
  ) +
  
  # Theme customization
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(size = 8, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(size = 8, color = "darkgrey", hjust = 1),
    legend.position = "top",
    legend.title = element_text(face = "bold", size = 11),
    legend.text = element_text(size = 9),
    legend.key = element_blank(),
    axis.text = element_text(size = 9),
    axis.title = element_text(size = 9),
    strip.text = element_text(size = 7, face = "bold"),
    panel.grid.major.y = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(20, 20, 20, 20)
  )

print(main_trends)
```

## 6. Data and output storage for easily reproduction

I will save the information to the right folder: 

- Raw Data
```{r}
write.csv(indicators_catalog, file.path(API_data, "indicators_catalog.csv")) # Save indicators catalog
write.csv(USMCA_indicators_raw, file.path(API_data, "USMCA_raw.csv"))
```

- Processed Data
```{r}
write.csv(findex_indicators_long, file.path(processed_data, "findex_indicators_long.csv"), row.names = FALSE) #findex
write.csv(USMCA_indicators, file.path(processed_data, "USMCA_indicators.csv"), row.names = FALSE) #USMCA
write.csv(all_indicators, file.path(processed_data, "all_indicators.csv"), row.names = FALSE) #Consolidated
```

- Outputs
```{r}
# Design a list and a loop for saving plots:
plots_to_save <- list(
  "outputs/global_inclusion_2017.png" = plot_2017,
  "outputs/global_inclusion_2021.png" = plot_2021,
  "outputs/borrowing_and_inflation.png" = borrow_inflation,
  "outputs/account_trend.png" = account_trend,
  "outputs/main_trends.png" = main_trends
)

# Save all plots
for (filename in names(plots_to_save)) {
  ggsave(plots_to_save[[filename]], filename = filename, width = 10, height = 7, dpi = 96, bg = "white") #dpi set to 96 for saving data memory and background white
}
```

*Note:* Some files were saved in previous sections because they contained raw data, making it easier to store them at that stage. For more details, please refer to the ReadMe file.

